### 2010-07-01 trunk/paper/IIWAS2010/content/related_work.tex
latest commit before added joirdan altered diagram

Several community QA systems have been developed, such as Yahoo! Answers \cite{yahooanswers2008}. These sites enlist a community of users to answer questions posed by other users.  This idea has led to socially integrated QA systems such as Aardvark \cite{vark2010}. In addition to having questions answered by users, Aardvark can contact potential question answerers through media beyond webpages (e.g. instant messages, mobile phones).  This allows for faster responses to questions.  Morpheus relies on both its user interface and previous similar questions answered by a community of users to give a fast reliable answer.

@@

The DBpedia\footnote{http://dbpedia.org} project is based on a community of contributors extracting semantic information from the Wikipedia and making it available on the Web.  Wikipedia semantics includes info-box templates, categorization information,  images, Geo-coordinates, links to external Web pages, disambiguation pages, and redirects between pages in Wiki markup form \cite{Bizer2009}.  DBpedia does not define any new relations between the Wikipedia categories.  Morpheus the DBpedia categories and its semantics to construct its ontology.
%  DBpedia represents the Wikipedia categories using skos:concepts and category relations using skos:broader \cite{Bizer2009}. In fact,  DBpedia does not define any new relations between the Wikipedia categories.  The Wikipedia categories are extracted directly from Wikipedia pages, and there  is no quality check on the resultant ontology. Similarly, additional ontological relations in DBpedia are generated from the info-box attributes and their values.

YAGO is a semi-automatically constructed ontology from the Wikipedia pages, info-boxes, categories and WordNet\footnote{http://wordnet.princeton.edu} synsets heterarchy \cite{Suchanek2009phd}. YAGO uses the Wikipedia page titles as its ontology individuals and categories as its ontology classes. YAGO uses only the nouns from WordNet and ignores the WordNet verbs and adjectives.  YAGO discovers connections between a WordNet synsets and a Wikipedia categories parsing the category names and matching the parsed category components with the WordNet synsets \cite{Suchanek2009phd}. Those Wikipedia categories having no WordNet match are ignored in the YAGO ontology.  The ontology's heterarchy is built using the \textit{hypernym} and \textit{hyponym} relations of the WordNet synsets.  Morpheus uses many of the same ideas made when using YAGO to develop its own ontology as explained in Section \ref{sec:ontology_corpora}.

>>>>

Several community QA systems have been developed, such as Yahoo! Answers \cite{yahooanswers2008}. These sites enlist a community of users to answer questions posed by other users.  This idea has led to socially integrated QA systems such as Aardvark \cite{vark2010}. In addition to having questions answered by users, Aardvark can contact potential question answerers through media beyond web pages (e.g. instant messages, mobile phones).  This allows for faster responses to questions.  Morpheus relies on both its user interface and referencing similar questions previously answered.

@@

The DBpedia\footnote{http://dbpedia.org} project is based on a community of contributors extracting semantic information from the Wikipedia and making this information available on the Web.  Wikipedia semantics includes info-box templates, categorization information,  images, geo-coordinates, links to external web pages, disambiguation pages, and redirects to pages in Wiki markup form \cite{Bizer2009}.  DBpedia does not define any new relations between the Wikipedia categories.

%Morpheus uses the DBpedia categories and its semantics to construct its ontology.
%  DBpedia represents the Wikipedia categories using skos:concepts and category relations using skos:broader \cite{Bizer2009}. In fact,  DBpedia does not define any new relations between the Wikipedia categories.  The Wikipedia categories are extracted directly from Wikipedia pages, and there  is no quality check on the resultant ontology. Similarly, additional ontological relations in DBpedia are generated from the info-box attributes and their values.

YAGO is a semi-automatically constructed ontology obtained from the Wikipedia pages, info-boxes, categories, and WordNet\footnote{http://wordnet.princeton.edu} synsets heterarchy \cite{Suchanek2009phd}. YAGO uses the Wikipedia page \emph{titles} as its ontology \emph{individuals} and \emph{categories} as its ontology \emph{classes}. YAGO uses only the nouns from WordNet and ignores the WordNet verbs and adjectives.  YAGO discovers connections between WordNet synsets and Wikipedia categories, parsing the category names and matching the parsed category components with the WordNet synsets. Those Wikipedia categories having no WordNet match are ignored in the YAGO ontology.  The ontology's heterarchy is built using the \textit{hypernym} and \textit{hyponym} relations of the WordNet synsets.

We use YAGO's principles to construct ontologies that provide similarity measures for answering questions within the same domain.  Thus far, these ontologies can be used to classifiy terms, however their classes do not always appropriately categorize query parameters.  Therefore, we are researching XXX in order to XXX.% TODO - fill in the XXX

%Morpheus uses many of the same ideas made when using YAGO to develop its own ontology as explained in Section \ref{sec:ontology_corpora}.

### 2010-07-01 trunk/paper/IIWAS2010/content/results_discussion.tex
latest commit before added joirdan altered diagram

First, we built an ontology for the automotive realm exploiting the Wikipedia pages, categories, and WordNet synsets. For each of the classes in the ontolgy we built corpora from the corresponding Wikipedia pages. Figure \ref{fig:automotive_ontology} shows a subsection of this ontology.

 describes the question answering process for the user query, ``A 1997 Toyota Camry V6 needs what size tires?''

%% TODO - need to refer back to the table which is figure one.
%The Morpheus NLP engine parses this query into the constructs: WH:what, Descriptive Information:1997 Toyota Camry V6, Asking For:size tires. From the descriptive info, we generate the N-grams or terms 1997, 1997 Toyota, 1997 Toyota Camry,Toyota,Toyota Camry,Toyota Camry V6, Camry, Camry V6, and V6. For each of the terms, we determined relevant categories (non-increasing order of relevance) from the ontology corpora. Table \ref{tbl:term_categories} shows the top categories and their probabilities for each of the query term.
In Table \ref{tbl:nlp_engine_parse} we show the data that is output by the Morpheus NLP engines parse of the query .  It extracts the WH question

\begin{table}[h]\footnotesize
	\begin{tabular}{|l|p{4.2cm}|}
		\hline 
		WH & what \\
		\hline 
		Descriptive Information & 1997 Toyota Camry V6 \\
		\hline 
		Asking for & size tires \\
		\hline 
		ngrams & 1997, 1997 Toyota, 1997 Toyota Camry, Toyota, Toyota Camry, Toyota Camry V6, Camry, Camry V6, V6 \\
		\hline
	\end{tabular}
	\caption{The output of NLP engine parse}
	\label{tbl:nlp_engine_parse} 
\end{table}

@@

\begin{tabular}{| p{3.5cm} | p{3cm} | r |}
\hline
Query & Tagged Classes & Score\\ \hline
A 1997 Toyota Camry V6 needs what size tires? & IC:Sedans, Cars, Engines, Manufactures OC:size tires & 0.91\\ \hline 
What is the tire size for a 1998 Sienna XLE Van? & IC:Vans, Manufactures OC: tire size & 0.72\\ \hline 
%Where can I buy a headlight for a Toyota Camry V6? & IC:Manufactures, Sedans, Engines OC: where & - \\ \hline 
%What is the cost of a Toyota Camry V6 muffler? &  IC:Manufactures, Sedans, Engines OC: the clost  & 0.97 \\ \hline
\end{tabular}

@@

Manually forming a corpus of wilipedia pages associated with a class is a cumbersome task. Topic Modeling provides a promising approach to identifying pages relevent to given class in a more automated manner.
%The Wikipedia page categories has limitations. Without supervision, the page categorization is often a challenging problem, because topics in the pages that belong to a realm (e.g. automotive) usually overlap. So we are evaluating the applicability of probabilistic topic models to web page categorization, and automatic generation of a realm-ontology based on the topics being extracted.

>>>>

First, we built an ontology for the vehicular realm exploiting the Wikipedia pages, categories, and WordNet synsets. For each of the classes in the ontolgy we built corpora from the corresponding Wikipedia pages. Figure \ref{fig:vehicular_ontology} shows a subsection of this ontology.

%describes the question answering process for the user query, ``A 1997 Toyota Camry V6 needs what size tires?''

%% TODO - need to refer back to the table which is figure one.
%The Morpheus NLP engine parses this query into the constructs: WH:what, Descriptive Information:1997 Toyota Camry V6, Asking For:size tires. From the descriptive info, we generate the N-grams or terms 1997, 1997 Toyota, 1997 Toyota Camry,Toyota,Toyota Camry,Toyota Camry V6, Camry, Camry V6, and V6. For each of the terms, we determined relevant categories (non-increasing order of relevance) from the ontology corpora. Table \ref{tbl:term_categories} shows the top categories and their probabilities for each of the query term.
In Table \ref{tbl:nlp_engine_parse} we show the data that is output by the Morpheus NLP engines parse of the query .  It extracts the WH question, that is, the the term that identifies the sentence as a question.  Next, the engine locates the descriptive information that can be used to describe the answer. Then the engine extracts the terms that describe what is being requested. Finally, the engine produces n-grams from the terms in the descriptive information sections.

Using the data in \ref{tbl:nlp_engine_parse} we determined relevant classes (non-increasing order of relevance) from the ontology given each n-gram. Table \ref{tbl:term_categories} shows the top classes and their probabilities for our example.

\begin{table}[h]\footnotesize
	\begin{tabular}{|l|p{4.2cm}|}
		\hline 
		WH & what \\
		\hline 
		Descriptive Information & 1997 Toyota Camry V6 \\
		\hline 
		Asking for & size tires \\
		\hline 
		n-grams & 1997, 1997 Toyota, 1997 Toyota Camry, Toyota, Toyota Camry, Toyota Camry V6, Camry, Camry V6, V6 \\
		\hline
	\end{tabular}
	\caption{The output of NLP engine parse}
	\label{tbl:nlp_engine_parse} 
\end{table}

@@

\begin{tabular}{| p{3.5cm} | p{3cm} | r |}
	\hline
	Query & Tagged Classes & Score\\ 
	\hline
	A 1997 Toyota Camry V6 needs what size tires? & inputs(Sedans, Cars, Engines, Manufactures) & 0.91\\ 
	\hline 
	What is the tire size for a 1998 Sienna XLE Van? & inputs(1998 Toyota Sienna XLE Van: Vans, Toyota : Manufacturers) & 0.72\\ 	\hline 
	Where can I buy an engine for a Toyota Camry V6? & inputs(Toyota Camry : Sedans, Toyota Camry V6 : Automobile_Engines, Toyota : Manufacturers) & 0.74 \\
	\hline 
%What is the cost of a Toyota Camry V6 muffler? &  IC:Manufactures, Sedans, Engines OC: the clost  & 0.97 \\ \hline
\end{tabular}

@@

Manually forming a corpus of wilipedia pages associated with a class is a cumbersome task. Topic Modeling provides a promising approach to identifying pages relevent to given class in a more automated manner.
%The Wikipedia page categories has limitations. Without supervision, the page categorization is often a challenging problem, because topics in the pages that belong to a realm (e.g. vehicular) usually overlap. So we are evaluating the applicability of probabilistic topic models to web page categorization, and automatic generation of a realm-ontology based on the topics being extracted.

### 2010-07-01 trunk/paper/IIWAS2010/content/system_arch.tex
latest commit before added joirdan altered diagram

Morpheus uses an ontology that contains classes of a particular realm of interest. Each leaf node in the ontology is associated with a corpus of words belonging to a class.  For example, we have constructed an automotive ontology containing classes relevant to the automotive realm. This ontology provides a structure for reference in the following sections.

@@

Morpheus uses the DBpedia categories, Wikipedia pages, and the WordNet synset heterarchy to find class-relevant web pages. This approach is motivated by YAGO \cite{Suchanek2009phd}. First a realm is mapped to a DBpedia category \cite{Bizer2009}. Then a \emph{Markov Blanket} \cite{PRIS} is created covering all neighboring categories and using the DBpedia ontology properties \emph{broader} and \emph{narrower}.

@@

In addition, we evaluate Latent Dirichlet Allocation (LDA) to identify latent topics of the documents in a corpus \cite{Blei2003latentdirichlet}.  LDA is Bayesian model represents a document in the corpus by distributions over topics, and a topic itself is a distribution over all  terms in the corpus.  For example, for the Wikipedia pages, the latent topics reflect the thematic structure of the pages. Thus, LDA can discover relevant topic proportions in a document using posterior inference \cite{Blei2003latentdirichlet}. Additionally, given a text document, using LDA we could tag related documents by matching the similarity over the estimated topic proportions, which is useful in the ontology and corpora building.  Due to its fully generative semantics, even at the level of documents, LDA could address the drawbacks (such as dimensionality and failure in finding the discriminative set of words for a document) of frequency based approaches such as TF-IDF, LSI, and pLSI.

@@

%% Was edited and shortened by Joir-dan and Wilson (2010-06-29)
The \emph{Query Resolution Recorder} (QRR) is an instrumented web browser that records the interactions of a path-finder answering a question. The path-finder also uses the tool to identify ontological categories associated with search terms. Morpheus stores the query together with its terms and categories as an SSQ (Figure \ref{fig:ssq_example}). Using the QRR, the path-finder is also able to identify where answers can be found within traversed pages.

%% Was added by Joir-dan and Wilson (2010-06-29)... but may need to be better formatted
\begin{figure}[t]
\label{fig:ssq_example}
("An  1997 Toyota Camry V6 has what size tires", 
     (1997: Date, Toyota: Manufacturer, Camry: Model, V6: Engine ),
     (tire: Part, size: Measurement))

%% Comment made by Joir-dan Gumbs (2010-06-29) 
%%Need to talk a bit about this SSQ Structure: Does anyone see anything wrong with this representation?
\caption{An example SSQ for the question: A 1997 Toyota Camry V6 has what size tires?}
\end{figure}

%% Edited by Joir-dan and Wilson (2010-06-29)
The \emph{Query Resolution Method} (QRM) is a data structure that models the
question answering process. A QRM represents a generalized executable realization of the search process that the path-finder followed. The QRM is able to reconstruct the page search path followed by the path-finder. Each QRM contains an \emph{ontological realm}, an SSQ, and information to support the query answering process. For each dynamic page, it stores the list of inputs to the URL such as form inputs and the possible reference outputs (highlighted portions of the web page).

@@

Queries inserted by path-followers take a different approach. The Morpheus search system parses, tags, and processes the queries in order to pull important n-grams. The system then assigns the most probable realm to the query using $p(realm|n-gram)$ priors calculated from realm-specific corpora. Once the realm is assigned, an ontology search is performed to assign classes to the n-grams. An SSQ is constructed and the system attempts to match this SSQ to a query path created by a path-finding user. Rather than matching n-grams, the goal is to match input and output classes, since the paths can potentially answer similar queries.

@@

To answer a user's query, a \textit{candidate
SSQ}, Morpheus finds similar SSQs that belong to QRMs in the Morpheus data store (a \textit{qualified SSQ}).  To determine SSQ similarity, we
consider the SSQ's realm, input terms, output terms, and their assigned classes.

@@

\noindent where $h$ is the height of the ontology tree.

@@

% Algorithm 1: figure  
\begin{figure}[t]
\centering
\includegraphics[width=85mm]{img/automotive_ontology.eps}
\caption{An abbreviated man-made automotive ontology - arrows represent inheritance relationships between the ontological classes.}
\label{fig:automotive_ontology}
\end{figure}

Suppose we want to find the category divergence between \textit{Sedans}
and \textit{Trucks} (Figure \ref{fig:automotive_ontology}). 
\textit{Sedans} is the subcategory of \textit{Cars}, and \textit{Trucks}
is the subcategory of \textit{Automobiles}. Therefore, the least common ancestor ($C$)
for these twoThis section presents ontology and copora, Query processing, ranking queries, and the query resolution mechanism. categories is \textit{Automobiles}. In addition, the tree height \textit{h = 4}.
So, the normalized divergence $cd$\textit{(Sedans, Trucks)} is \textit{7/12}
that is calculated from \textit{d(Sedans, Root) = 4}, \textit{d(Sedans, Automobiles) = 2}, and \textit{d(Trucks, Automobiles) = 1}.

@@

 The evaluation of this script by the QRE simulates the behavior of a human browsing the web, that is, clicking buttons to follow links, submitting forms, and highlighting of data to form a text answer.  The QRE assumes that because of the auto generated nature of deep web pages, the location of answers are the same irrespective of page changes.  It uses the relative XPath location to the answer node on HTML pages as described in \cite{Badica06}.

>>>>

Morpheus uses an ontology that contains classes of a particular realm of interest. Each leaf node in the ontology is associated with a corpus of words belonging to a class.  For example, we have constructed an vehicular ontology containing classes relevant to the vehicular realm. This ontology provides a structure for reference in the following sections.

@@

Morpheus uses the DBpedia categories, Wikipedia pages, and the WordNet synset heterarchy to find class-relevant web pages. First a realm is mapped to a DBpedia category \cite{Bizer2009}. Then a \emph{Markov Blanket} \cite{PRIS} is created covering all neighboring categories and using the DBpedia ontology properties \emph{broader} and \emph{narrower}.

@@

In addition, we evaluate Latent Dirichlet Allocation (LDA) to identify latent topics of the documents in a corpus \cite{Blei2003latentdirichlet}.  LDA is Bayesian model that represents a document in the corpus by distributions over topics, and a topic itself is a distribution over all  terms in the corpus.  For example, for the Wikipedia pages, the latent topics reflect the thematic structure of the pages. Thus, LDA can discover relevant topic proportions in a document using posterior inference \cite{Blei2003latentdirichlet}. Given a text document, we tag related documents by matching their similarity over the estimated topic proportions, assisting in ontology and corpora construction.  Due to its fully generative semantics, this usage of LDA could address drawbacks of frequency based approaches (e.g. TF-IDF, LSI, and pLSI) such as dimensionality and failure to find the discriminative set of words for a document. %of frequency based approaches.% such as TF-IDF, LSI, and pLSI.

@@

The \emph{Query Resolution Recorder} (QRR) is an instrumented web browser that records the interactions of a path-finder answering a question. The path-finder also uses the tool to identify ontological categories associated with search terms. Morpheus stores the query together with its terms and categories as an SSQ.  Table \ref{tab:ssq_example} is an example showing the SSQ model of the query ``A 1997 Toyota Camry V6 has what size tires?". The SSQ in Table \ref{tab:ssq_example} is said to be \emph{qualified} because the classes associated with its terms have been identified. Using the QRR, the path-finder is also able to identify where answers can be found within traversed pages.

%	\begin{figure}[t]
%\label{fig:ssq_example}
%("An  1997 Toyota Camry V6 has what size tires", 
%     (1997: Date, Toyota: Manufacturer, Camry: Model, V6: Engine ),
%     (tire: Part, size: Measurement))
%\caption{An example SSQ for the question: A 1997 Toyota Camry V6 has what size tires?}
%\end{figure}

%Query & \multicolumn{9}{|c|}{A 1997 Toyota Camry V6 has what size tires?}\\

%% TODO - we need the table here
\begin{table}\footnotesize
\scalebox{0.76}{
	\rowcolors{1}{white}{white}
	\begin{tabular}{|r|l|l|l|l|l|l|l|l|}
		\hline
		\textbf{Terms} & 1997 & Toyota & Camry & V6 & size & tires \\
		\hline
		\textbf{Input} & Date & Manufacturer & Model & Engine & &  \\
		\hline
		\textbf{Output} & & & & & Measurement & Part \\
		\hline
	\end{tabular}
	}
	\caption{Example SSQ Model.}
	\label{tab:ssq_example}
\end{table}

The \emph{Query Resolution Method} (QRM) is a data structure that models the question answering process. A QRM represents a generalized executable realization of the search process that the path-finder followed. The QRM is able to reconstruct the page search path followed by the path-finder. Each QRM contains an \emph{ontological realm}, an SSQ, and information to support the query answering process. For each dynamic page, the QRM contains a list of inputs to the URL such as form inputs and the possible reference outputs.

@@

Queries inserted by path-followers take a different approach. The Morpheus search process parses and tags queries in order to record important terms. The system assigns the most probable realm given the terms in the query as calculated from realm-specific copora. Once the realm is assigned, an ontology search is performed to assign classes to the terms. An SSQ is constructed and the system attempts to match this candidate SSQ to existing QRM SSQs. Rather than matching exact query terms, the system matches input and output classes, because a QRM can potentially answer similar queries.

@@

To answer a user's query, a candidate SSQ, Morpheus finds similar qualified SSQs that are associated with QRMs in the Morpheus data store.  To determine SSQ similarity, we consider the SSQ's realm, input terms, output terms, and their assigned classes.

@@

\noindent where $h$ is the longest path in the ontology class heterarchy.

@@

% Algorithm 1: figure  
\begin{figure}[t]
	\centering
%\includegraphics[width=85mm]{../img/vehicular_ontology.eps}
\includegraphics[width=85mm]{vehicular_ontology.eps}
\caption{Abbreviated Vehicular Ontology}% - arrows represent inheritance relationships between the ontological classes.}
\label{fig:vehicular_ontology}
\end{figure}

Suppose we want to find the category divergence between \textit{Sedan} and \textit{Bus} Figure \ref{fig:vehicular_ontology}. \textit{Sedan} is the subcategory of \textit{Sedan}, and \textit{Bus} is the subcategory of \textit{Automobile}. Therefore, the least common ancestor ($C$) for these two. This section presents ontology and copora, query processing, ranking queries, and the query resolution mechanism. categories is \textit{Automobile}. In addition, the tree height \textit{h = 4}. So, the normalized divergence $cd$\textit{(Sedan, Bus)} is \textit{7/12} that is calculated from \textit{d(Sedan, Root) = 4}, \textit{d(Sedan, Automobile) = 2}, and \textit{d(Bus, Automobile) = 1}.

@@

The Morpheus \emph{Query Executor} (QE) evaluates a script of the query resolving process.  It simulates a human clicking buttons to follow links, submitting forms, and highlighting of data to form a textual answer.  The QE assumes that because of the auto generated nature of deep web pages, the location of answers are the same irrespective of page changes.  It uses the relative XPath location to the answer node on HTML pages as described in \cite{Badica06}.
